UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-09-05 ms/main.tex
tweak wording on clustering

By patching together many local genetic neighbourhoods, UMAP visualizations preserve aspects of both local and global structure. Methods such as t-SNE and LargeVis\cite{tang2016visualizing} are similarly motivated, with key differences in how they measure neighbours and which cost functions are used in optimizing low-dimensional embeddings: pairwise Gaussian and t-distributed approximations and the KL divergence in t-SNE; approximate nearest neighbours and a likelihood function in LargeVis; and fuzzy simplicial set membership and the cross-entropy in UMAP. Building the global structure in t-SNE requires tuning the \textit{perplexity} parameter, which in practice will need to be different for varying cluster sizes\cite{wattenberg2016how}, while UMAP merges discrete simplicial sets into a coherent global structure. Further practical and theoretical comparisons can be found in~\cite{2018arXivUMAP}.

>>>>

By patching together many local genetic neighbourhoods, UMAP visualizations preserve aspects of both local and global structure. Methods such as t-SNE and LargeVis\cite{tang2016visualizing} are similarly motivated, with key differences in how they measure neighbours and which cost functions are used in optimizing low-dimensional embeddings: pairwise Gaussian and t-distributed approximations and the KL divergence in t-SNE; approximate nearest neighbours and a likelihood function in LargeVis; and fuzzy simplicial set membership and the cross-entropy in UMAP. Further practical and theoretical comparisons can be found in~\cite{2018arXivUMAP}. Hierarchical clustering of networks has also been successful in detecting fine-scale population structure, though it was instead motivated by partitioning IBD data into subnetworks to then find stable subclusters\cite{han2017clustering}; in practice, one could carry this out by detecting clusters in data that has been pre-processed using UMAP as laid out in \cite{howtoUMAP}.
